Some figures to illustrate basic ideas in FDR

### 2016-12-12 analysis/referee.response.Rmd

---
title: "Response to referees"
author: "Matthew Stephens"
date: 2016-03-21
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk_options, include=FALSE}
source("chunk-options.R")
```

## Referee 1

@@

I agree with a lot of what the referee says, although there are some points of 
disagreement. The most fundamental disagreement regards Figure 1. This Figure 
shows not the detectable effects (as the referee believed) but the decomposition
of *all* effects into null and alternative components. Since this seems to be an 
important misunderstanding I have revised the text to try to make it clearer how
this figure was derived. Specifically the revised text says:

"We used each of the methods qvalue, locfdr, mixfdr and ash to decompose the $z$ 
scores ($z_j = \hat{\beta}_j$), or their corresponding p values, into null and 
alternative components. Here we are using the fact that these methods all 
provide an estimate of the lfdr for each observation, which implies such a 
decomposition; specifically the average lfdr within each histogram bin estimates
the fraction of observations in that bin that come from the null vs the 
alternative component.”

@@

Since the referee objected to the presentation of the ZA and UA as opposing 
assumptions I have removed all explicit reference to the ZA. (This also
solves the issue that I had defined the ZA differently than Efron’s original 
definition — see below — which may have created confusion.) Instead, as the 
referee suggested, I have presented the UA as an additional assumption “not 
made by other methods”, and - as the editor suggested — I have focussed on 
contrasting the behavior of ash with the behavior of the other methods (which, 
as Figure 1 demonstrates, is to create a hole at 0 in the distribution of the 
alternative Z scores) rather than contrasting their assumptions.

However, it seems unsatisfactory to merely point out this behavioral difference
between methods without giving any explanation for the reason behind it. So I 
have provided my best attempt to give the reason. Essentially the reason is 
that the existing methods (locfdr and qvalue), when estimating $\pi_0$, make 
the assumption that *all* p values near 1 are null, or *all* z scores near 0 
are null. (This was how I defined the ZA in the original paper; however this 
is stronger than Efron’s original statement of the ZA, which the referee - 
understandably - uses in his/her report, that ``most” z scores near 0 are null. 
So I don’t refer to this as the ZA any more.) This assumption necessarily 
creates a hole in the implied distribution of z scores under the alternative,
and explains the behavior.

To underscore that this behavior is general, and not just a feature of the 
particular simulation I did in Figure 1, I provide results using qvalue under
the setting suggested by the referee at; see  (http://stephenslab.github.io/ash/analysis/referee_uaza.html).

### Minor Concerns

@@

The referee says: "it would be interesting to include distributions that 
favor other models, such as a bimodal distribution with no mass at zero, an 
assymmetrical distribution,. . . The unimodality is a strong assumption so it 
would be good to see how anti-conservative the results would get under 
different alternative distributions."

I found this puzzling since both an assymetrical distribution and a strongly 
bimodal distribution are already included in the simulation results (Figure 2, 
Figure 3, Table 1). I believe these existing results address the referee comment. 
I also want to emphasise that the unimodal assumption is actually less strong 
than the assumptions usually made in variable selection for regression, which 
is directly analogous to the fdr context. For example, it is common to assume 
that effects are a mixture of a point mass at zero and a single normal, which 
is considerably more restrictive than the general unimodal assumption.

However, to help reassure the referee, I have added another example to 
illustrate an assymetric situation where the unimodal-at-zero assumption is 
badly contradicted 
[here](http://stephenslab.github.io/ash/analysis/efron.fcr.html). As
the referee will see, the resulting estimates of the lfdr from ashr are
reasonably close to the true values (and also to those from locfdr).

### Minor Concern

*Is there a way to model the distribution as a mixture of both uniforms and 
normals? If not: why?*

@@

This is straightforward in principle. However, it would require non-trivial 
code modifications because of the way I implemented the mixture class (assuming 
that all components  come from the same family.) Another way to achieve an 
assymetric distribution without the "hard" tails of the uniform might be to use 
a mixture of truncated normals. I have a student investigating this option. 
However, the uniform mixture has the advantage of allowing not only the 
assymetry, but also the $t$ likelihood. Although the paper does not focus on 
this, I believe this is likely to be an important issue in practice. So I 
suspect that the mixture of uniforms is often going to be the method of choice 
in practice.

*How would the model perform with moderate correlations?*

There seems to be no reason to think that moderate correlations among tests 
will have a substantial detrimental effect. But I do not want to focus on this 
because I think it will give the reader the wrong message: I think that users 
of both ashr and other fdr methods should very much worry about correlations 
in practice (which I believe will often be large, and not moderate).

*In equation 3, the null component (I think) is represented as delta0, but 
delta0 hasn’t been defined. What is assumed for the distribution of delta0?*

@@

Thank you for noticing this. I now define $\delta_0$ (which represents a point 
mass on 0).

*Figure 3: the truth is hard to distinguish from the different methods. And in
general this figure is hard to read.*

@@

I agree. I have completely rethought this figure, which now has fewer, larger,
panels, and compares just two  methods (ash.hu and the NPMLE) against the truth 
on a few examples. The less important point - the effect of the penalty term on 
pi0 -is now illustrated in an additional figure included in the appendix.

*I particularly like the idea that the ordering of the tests differs from the
classical p-values. Which influence does this have on the balance between
sensitivity and specificity (for example with ROC curves)? It should change, 
but does it get better?*

@@

Thanks for the question. This is now addressed in the new Figure 4d.

### Typos and grammar

*page 14, line 20. The first mentioned pi0 and lfdr should have a hat.*

### 2016-12-16 analysis/makefig_FDReg.Rmd

```{r chunk-options, include=FALSE}
library(ashr)
library(qvalue)
library(locfdr)
library(mixfdr)
source("chunk-options.R")
source("../R/plot_FDReg_hist.R")
source("../R/nullalthist.R")
```

>>>>

```{r chunk_options, include=FALSE}
library(ashr)
library(qvalue)
library(locfdr)
library(mixfdr)
source("chunk-options.R")
source("../R/plot_FDReg_hist.R")
source("../R/nullalthist.R")
```
